Co-citations in context

### 2019-05-20 cocit-r.tex
renamed the file back to original

\begin{abstract}
Citation analysis of the scientific literature has been used to study and define disciplinary boundaries, to trace the dissemination of knowledge, and to estimate impact. Co-citation, the frequency with which pairs of publications are cited, provides insight into how documents relate to each other and across fields. Co-citation analysis has been used to characterize combinations of prior work as conventional or innovative and to derive features of highly cited publications. Given the organization of science into disciplines, a key question is the sensitivity of such analyses to frame of reference. Our study examines this question using  semantically-themed citation networks. We observe that trends reported to be true across the scientific literature do not hold for focused citation networks, and conclude that co-citation analysis requires a contextual perspective.
\end{abstract}

@@

In an intriguing study, co-citation analysis of 17.9 million articles and their cited references from the Web of Science (WoS) was used to document novel combinations of prior research and to characterize highly cited papers~\citep{uzzi_atypical_2013}. In this study, pairwise combinations of the references in each article were generated to create a multiset of journal-pairs. Observed frequencies of these journal pairs across the dataset were computed and normalized to expected values generated by Monte Carlo simulations that shuffled cited references in the context of a random graph model. The resultant normalized, journal pair frequencies were termed \emph{z-scores} (Materials and Methods). Thus, every article was associated with multiple z-scores. For each article, positional statistics of these z-scores were calculated to describe conventionality;(high conventionality (HC) if the median z-score for an article was greater than the median of median z-scores of all articles and low conventionality for the inverse (LC). Similarly with novelty;  high novelty (HN) if the tenth percentile of z-scores for an article was less than zero and low novelty (LN) for the inverse. Accordingly, each article was labeled with respect to conventionality and novelty, e.g, HCLN, with all four combinations being possible. The finding of considerable interest was that HCHN articles were twice as likely to be found in highly cited papers~\citep{uzzi_atypical_2013} suggesting that novel ideas flavoring a base of conventional thought were a recipe for impact. The authors examined subsets of publications grouped by disciplined and asserted universality of this feature.

Key to the Uzzi study, however, is the random graph approach and the mechanism of citation shuffling employed. Random substitutions are made under conditions that preserve the number of publications in each year of the dataset as well as the number of references cited by each of these publications. However, the approach inadequately accounts for observed citation behavior~\citep{wallace_lariviere_gingras_2012,moed_measuring_2010,klavans_research_2017,garfield_1979} since (i) random substitutions are selected with equal probability irrespective of disciplinary origin (ii) replacements are drawn from the set of eligible references rather than the multi-set that accounts for the frequency with which a reference is cited. In illustration, a musicology reference from can replace a quantum physics reference during the simulation process. Similarly, a reference cited over 100 times in a given year is selected with the same probability as a reference cited only once. Accordingly, the expected value calculations generated by simulations and used in normalizing observed journal-pair-frequencies can be reasonably questioned on grounds of model mis-specification.

A follow-up study using a smaller dataset of 12 million articles, also suggests that Uzzi and colleagues do not fully account for disciplinary and journal effects~\citep{boyack_vs_uzzi_2014}. While the two studies are conceptually similar, the authors of the latter study used Scopus data and the K50 measurement, a derivative of the $\chi^2$ method, rather than Monte Carlo simulations for normalizing observed to expected journal-pair frequencies. Boyack and Klavans note that ``only 64.4\%  of  243  WoS  subject  categories'' in the Uzzi study met the criterion of having the highest probability of hit papers in the HCHN category. Further, they observed that journals vary widely in terms of size and influence and that 20 journals accounted for nearly 15\% of co-citations in their measurements. Lastly, that three multidisciplinary journals accounted for 9.4\% of all atypical combinations, suggesting strong effects from both disciplines and journals.

Despite different methods used to generate expected values, expected journal-pair frequencies were generated without disciplinary constraints in both these prior studies~\citep{uzzi_atypical_2013,boyack_vs_uzzi_2014}, after which subsets of publications were examined for disciplinary effects. To avoid the problem of discipline-insensitive substitutions, we chose instead to first construct semantically related sets of documents (disciplinary networks) and then measure observed and expected frequencies within these networks. Accordingly, we used keyword searches of the scientific literature to create three citation networks themed around broad search terms. For each of these, we calculated expected values using an efficient simulation algorithm that randomly permuted references grouped by year of publication in a random graph model that accounted for existing citation frequencies in these networks. We hypothesized that our approach would  reduce model misspecification and better simulate citation practice, in turn leading to different and more relevant conclusions.  Our study on these semantically-themed citation networks reveals significantly different patterns of conventionality and novelty between  citation networks and disciplines that challenges the conclusion of universality in HCHN articles have a greater probability of being represented in highly cited publications. Instead, we conclude that conventional thought as commonly defined in our study, Uzzi, and Boyack is more likely to drive higher citations.. \emph{these are just filler conclusions that must be carefully rewritten and added to}.

>>>>

\begin{abstract}
Citation analysis of the scientific literature has been used to study and define disciplinary boundaries, to trace the dissemination of knowledge, and to estimate impact. Co-citation, the frequency with which pairs of publications are cited, provides insight into how documents relate to each other and across fields. Co-citation analysis has been used to characterize combinations of prior work as conventional or innovative and to deriveå features of highly cited publications. Given the organization of science into disciplines, a key question is the sensitivity of such analyses to frame of reference. Our study examines this question using  semantically-themed citation networks. We observe that trends reported to be true across the scientific literature do not hold for focused citation networks, and conclude that co-citation analysis requires a contextual perspective.
\end{abstract}

@@

\cite{uzzi_atypical_2013} used a novel approach for co-citation analysis of 17.9 million articles and their cited references from the Web of Science (WoS)  to characterize the highly cited papers according to both novel and conventional combinations of prior research. %~\citep{uzzi_atypical_2013}. 
%In this study, pairwise combinations of the references in each article were generated to create a multiset of journal-pairs. 
%They explored the use of a random graph model, in which 
Observed frequencies of these journal pairs across the dataset were computed and compared to expected values generated by Monte Carlo simulations that shuffled cited references in the context of a random graph model, allowing substitutions of references by any other reference published in the same year and appearing in WoS.
By normalizing the differences in these frequencies, journal pairs were assigned \emph{z-scores} (Materials and Methods). 
Thus, every article was associated with multiple z-scores. 
For each article, positional statistics of these z-scores were calculated to describe conventionality: high conventionality (HC) if the median z-score for an article was greater than the median of median z-scores of all articles and low conventionality for the inverse (LC). 
Similarly,  an article was deemed to have high novelty (HN) if the tenth percentile of its z-scores  was negative and low novelty (LN) for the inverse. Accordingly, each article was labeled with respect to conventionality and novelty, e.g, HCLN (denoting that the article has high conventionality and low novelty), with all four combinations being possible. 
The finding of considerable interest was that HCHN articles were twice as likely to be found in highly cited papers~\citep{uzzi_atypical_2013} suggesting that novel ideas flavoring a base of conventional thought were a recipe for impact. 
The authors examined subsets of publications grouped by disciplined and asserted universality of this feature.

Key to the Uzzi {\em et al.} study, however, is the random graph approach and the mechanism of citation shuffling employed. 
Their random graph model produces
 substitutions of references by other references   that preserve the number of publications in each year of the dataset as well as the number of references cited by each of these publications. 
 However, the approach also assumes that all substitutions are equiprobable (i.e., a substitution of a reference in physics by a reference in musicology has the same probability under their model as a substitution of the physics reference by another physics reference).
 This assumption does not account %to replace a reference from the same year, with equal probability inadequately accounts 
 for observed citation behavior~\citep{wallace_lariviere_gingras_2012,moed_measuring_2010,klavans_research_2017,garfield_1979}, and clearly is inconsistent with author behavior (i.e., in practice, an author of a paper in evolutionary biology is more likely to cite a biology paper than a paper in art history). 
% Furthermore, the citation frequency of the reference does not impact the probability of selecting the
% reference, which also is inconsistent with author behavior. 
  %since (i) random substitutions are selected with equal probability irrespective of disciplinary origin (ii) replacements are drawn from the set of eligible references rather than the multi-set that accounts for the frequency with which a reference is cited. In illustration, a musicology reference from can replace a quantum physics reference during the simulation process. 
 In addition, under the Uzzi {\em et al.} random model,  a reference cited over 100 times in a given year is selected with the same probability as a reference cited only once. 
 Accordingly, the expected value calculations generated by simulations and used in normalizing observed journal-pair-frequencies can be reasonably questioned on grounds of model mis-specification.

A follow-up study using a smaller dataset of 12 million articles, also suggests that Uzzi and colleagues do not fully account for disciplinary and journal effects~\citep{boyack_vs_uzzi_2014}. While the two studies are conceptually similar, the authors of the latter study used Scopus data and the K50 measurement, a derivative of the $\chi^2$ method, rather than Monte Carlo simulations for normalizing observed to expected journal-pair frequencies. Boyack and Klavans note that ``only 64.4\%  of  243  WoS  subject  categories'' in the Uzzi study met the criterion of having the highest probability of hit papers in the HCHN category. Further, they observed that journals vary widely in terms of size and influence and that 20 journals accounted for nearly 15\% of co-citations in their measurements. Lastly, that three multidisciplinary journals accounted for 9.4\% of all atypical combinations, suggesting strong effects from both disciplines and journals. 
Despite different methods used to generate expected values, both studies generated expected journal-pair frequencies without disciplinary considerations~\citep{uzzi_atypical_2013,boyack_vs_uzzi_2014}, and only after this were subsets of publications  examined for disciplinary effects.

To avoid the problem of discipline-insensitive substitutions, we chose instead to first construct semantically related sets of documents (disciplinary networks) and then measure observed and expected frequencies within these networks. Accordingly, we used keyword searches of the scientific literature to create three citation networks themed around broad search terms. For each of these, we calculated expected values using an efficient simulation algorithm that randomly permuted references grouped by year of publication in a random graph model that accounted for existing citation frequencies in these networks. We hypothesized that our approach would  reduce model misspecification and better simulate citation practice, in turn leading to different and more relevant conclusions.  Our study on these semantically-themed citation networks reveals significantly different patterns of conventionality and novelty between  citation networks and disciplines that challenges the conclusion of universality in HCHN articles have a greater probability of being represented in highly cited publications. Instead, we conclude that conventional thought as commonly defined in our study, Uzzi, and Boyack is more likely to drive higher citations.. \emph{these are just filler conclusions that must be carefully rewritten and added to}.
